---
title: "Puppy Raiser Questionnaire vs Outcome (Success, Fail)"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_questionnaires")
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggpubr")
library(DescTools)
library(ggplot2)
library(cowplot)
library(dplyr)
library(plyr)
library(tidyverse)
library(rstatix)
library(ggpubr)
```

## Introduction 
This document tests the hypothesis that there is a difference in behaviours reported in the standardised questionnaire Canine Behavioural Assessement & Research Questionnaire (C-BARQ) considering dog's training outcome.

## Data Exploration
### Columns: Variables
Below are the questionnaire items (variables) that were scored by the Puppy Raise in a scale from 'Never', 'Seldom', 'Sometimes', 'Usually', 'Always'. 

```{r, echo = FALSE}
prq = read.csv('data//2022-05-10-PRQ_C-BARQ.csv')
vars = read.csv("data//questionnaire-variables-puppyraisers.csv",stringsAsFactors=FALSE)
# print the number of variables per data type
# print(table(vars$Type))

# Adding a column to df vars for Test 
data = prq %>% select(contains(c("X\\d", "F_", "Outcome")))
colnames(data)
```


### Rows: Dogs
The number of dogs in the Puppy Raiser Questionnaire dataset. Selection criteria: dogs with a defined training outcome.
```{r, echo = FALSE}
#cat('Number of dogs:', length(prq$Name))
prq = prq %>% distinct(Code, .keep_all = TRUE)
cat('Training Outcome:')
table(prq$Outcome)
```
Analysing categorical demographic data: Sex, Breed
```{r, echo = FALSE}
table(prq$Sex)
table(prq$Breed)
```
Analysing age at assessment:
```{r, echo = FALSE}
prq$Age.at.Assessment <- prq$Timestamp - prq$DOB
cat('Age at Assessment: Mean', mean(prq$Age.at.Assessment)/30.417, 
            'Standard Deviation', sd(prq$Age.at.Assessment)/30.417)
```



## Data Analysis
Logistic Regression one-way analysis of variance test was used to investigate whether there was a statistically significant difference between the behaviours reported by Puppy Raisers and the Training Outcome of the dogs.
C-BARQ Correlation with Training Outcome. 

```{r, echo = FALSE}

```

