---
title: "Dog Trainer Questionnaire vs Outcome (Fail, Success) "
author: "Marinara Marcato"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_questionnaires")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
#install.packages("ggpubr")

#install.packages("ggpubr")
library(DescTools)
# plot
library(ggplot2)
library(ggpubr)
library(cowplot)
# datasets
library(plyr)
library(dplyr)
library(tidyverse)
library(reshape2) # remove?
# machine learning/stats
library(recipes)
library(caret)
library(mlbench) # remove?

```

# Introduction 
In order to objectively evaluate trainee guide dogs personality, their dog trainers filled out the standardised questionnaire Monash Canine Personality Questionnaire - Revised (MCPQ-R) after 10 weeks of training. 
This document shows the data analysis carried out to investigate the association between the ratings given by trainers and the dog’s training outcome. 
Statistical methods will be used to test the hypothesis that there is a relationship between personality ratings and training outcome (Success, Fail).

# Data Exploration

Importing data and converting variables to adequate data types.
```{r, echo = FALSE}
dtq = read.csv('data//2022-06-27-DTQ_MCPQ-R.csv', stringsAsFactors=TRUE)
# colnames(dtq)
# converting date types
dtq$Timestamp = as.Date(dtq$Timestamp, format= "%Y-%m-%d")
dtq$DOB = as.Date(dtq$DOB, format= "%Y-%m-%d")
dtq$DOA = as.Date(dtq$DOA, format= "%Y-%m-%d")
dtq$End.Date = as.Date(dtq$End.Date, format= "%Y-%m-%d")
dtq$Duration = as.numeric(gsub(" .*$", "", dtq$Duration))
```
## Demographics
The number of dogs in the Dog Trainer Questionnaire dataset. Selection criteria: dogs who successed training (assistance and guide dogs) and dogs withdrawn for behavioural reasons.
```{r, echo = FALSE}
cat('Number of dogs:', length(dtq$Code))
cat('Training Outcome:')
table(dtq$Outcome)
```

Analysing categorical demographic data: Sex, Breed. There was only one German Shepherd and two Golden Doodle dogs in the sample, their breeds were relabeled as "Other" for the data analysis.
```{r, echo = FALSE}
print(table(dtq$Sex))
print("Original classes")
table(dtq$Breed)
# merging breed categories
levels(dtq$Breed)[levels(dtq$Breed) == "LRx"] <- "LRxGR"
levels(dtq$Breed)[levels(dtq$Breed) =="GS" | levels(dtq$Breed) =="GRxPoodle"] <- "Other"
print("Processed classes")
table(dtq$Breed)
```

Analysing age at arrival to the training centre and age at assessment when the questionnaire was completed:

```{r, echo = FALSE}
n <- dim(dtq)[1]

dtq$Age.at.Arrival <- dtq$DOA - dtq$DOB
mean <- mean(dtq$Age.at.Arrival)
std <- sd(dtq$Age.at.Arrival)
margin <- qt(0.975,df=n-1)*sd(std)/sqrt(n)

cat('Age at Arrival: Mean', round(mean/30.417, 2), 
            'Standard Deviation', round(std/30.417, 2))
            # 'Confidence Interval', round((mean-margin)/30.417, 2), round((mean+margin)/30.417, 2)


dtq$Age.at.Assessment <- dtq$Timestamp - dtq$DOB
cat('Age at Assessment: Mean', round(mean(dtq$Age.at.Assessment)/30.417,2), 
            'Standard Deviation', round(sd(dtq$Age.at.Assessment)/30.417, 2))


dtq$Duration.at.Assessment <- dtq$Timestamp - dtq$DOA
cat('Duration at Assessment: Mean', round(mean(dtq$Duration.at.Assessment)/30.417,2), 
            'Standard Deviation', round(sd(dtq$Duration.at.Assessment)/30.417, 2))
summary(as.integer(dtq$Duration.at.Assessment)/7)
hist(as.integer(dtq$Duration.at.Assessment)/7)
```

Calculate statistics of duration of training for the dogs that were withdrawn from training. 
```{r, echo = FALSE}
# Duration of training before withdrawal in weeks
# dtq %>% arrange(Duration) %>% select(Duration, Timestamp, DOA, Name)
duration <- dtq %>% filter(Outcome == "Fail") %>% select(Duration)/7
print('Duration of Training in weeks')
summary(duration)

h <- ggplot(duration, aes(x=Duration)) +
 geom_histogram(binwidth = 1) +
 xlab("Duration (Weeks)") +
 ylab("Number of Dogs") + theme_bw()
 
h
suppressMessages(ggsave("results/dtq/duration-histogram.png",h))
```

## Descriptive Statistics
The MCPQ-R contains 26 items which were scored by the Trainer in a scale from 1 to 6.
The questionnaire data are kept as integers, rather than being converted to factors, so the information about the ordering is kept.
The descriptive statistics of the questionnaire data shown below are saved as a csv.  

<!-- 
Other potential variables for inclusion in the model:
"PR.Sup" -> Alison = 5, Catherine = 6, DSM = 3, Frances = 11, Graham = 24, Mags = 3, Rose = 23, UKPR = 11
"Source" -> IGDB = 60, AADI =3, Cesecah = 1, GDBUK = 16, Private breeder = 8, PRV breeder = 1 -->

```{r, echo = FALSE}
# selecting features for modelling
data = dtq %>% select(contains(c("Extraversion", "Motivation", "Training",
 "Amicability", "Neuroticism", "Sex", "Breed", "Outcome")))  %>% select(!contains("Comments"))
data$Outcome <- relevel(data$Outcome, "Success")
cat("Dimension of the dataset including the dependent variable: ", dim(data))
# str(data)
# calculate descriptive statistics on for the numeric factors
stats <- data.frame(do.call(rbind, lapply(data %>% select(where(is.numeric)), summary)))
print(stats)
# lapply(data %>% select(where(is.factor)), summary)
# save descriptive statistics of the dataset to csv
write.csv(stats, "results/dtq/descriptive-statistics.csv") 
```

## Univariate Logistic Regression
Testing if variables have a statistically significant association with Outcome (Fail, Success). 
Univariate Logistic Regression models use each feature individually to predict the outcome.
Features with a p_value < 0.2 were kept for the multicollinearity analysis and further multivariate logistic regression model.

```{r, echo = FALSE}
# univariate logistic regression using one predictor and outcome
lr_models <- lapply(data %>% select(-Outcome), 
        function(x) summary(glm(formula = Outcome ~ x, 
        data = data, 
        family = binomial(link = logit))))
lr_result <- lapply(lr_models, function(x) round(c(coef(x), x$deviance),4))

## univariate models with 3 dfs (Breed)
print(lr_models[28])
lr_breeds <- data.frame(transpose(lr_result[28]))
colnames(lr_breeds) <- c("estimate_0", "estimate_1", "estimate_2", 
                    "se_0", "se_1", "se_2", "z_value_0", "z_value_1", "z_value_2", 
                    "p_value_0", "p_value_1", "p_value_2", "deviance")
# print(unique(data$Breed))
cat('Difference between Other and LR: \t', lr_breeds$p_value_1, '\t -> not significant')
cat('Difference between Other and LRxGR: \t', lr_breeds$p_value_2, '\t -> not significant')

## univariate models with up to 2 dfs
lr_results <- data.frame(do.call(rbind,lr_result[1:27]))
colnames(lr_results) <- c("estimate_0", "estimate_1", 
                    "se_0", "se_1", "z_value_0", "z_value_1",
                    "p_value_0", "p_value_1", "deviance")
lr_results %>% filter(p_value_1 < 0.05) %>% select(estimate_1, p_value_1)

# saving univariate logistic regression results to csv
write.csv(lr_results, "results/dtq/univariate-logistic-regression.csv")
```

```{r, echo = FALSE}
# select questionnaire items based on the p values
feat_lr <- lr_results %>% filter(p_value_1 < 0.05) %>% rownames()
data_lr <- data %>% select(all_of(feat_lr), Outcome)
```

## Multicollinearity 
The Variance Inflation Factor (VIF) was calculate to assess multicollinearity. 
VIF = 1 no correlation, 5 < VIF < 10 moderate correlation, VIF > 10 high multicollinearity. 
VIF larger than 5 or 10 is large and indicate a high level of multicollinearity and should be further processed.

```{r, echo = FALSE}
# including all features selected by lr test
model_lr <- glm(Outcome ~ ., data=data_lr, family = binomial(link = logit)) 
summary(model_lr)
print(VIF(model_lr))
```

## Visualization
Plotting features selected considering univariate logistic regression pvalue results.

```{r, echo = FALSE}
plot_items <- function(var){
    
    plot1 <- ggplot(data_lr, aes_string( x = var, fill = 'Outcome', width = 1)) + 
            geom_bar(position = "dodge") +
            theme(legend.position = c(0.84,0.88)) 
            # scale_x_discrete("Rating", limits = c(0:4), breaks = c(0:4), labels = c(0:4)) +
            # expand_limits(x = c(0,4))

    plot2 <- ggplot(data_lr, aes_string( x = 'Outcome', y = var)) + 
            geom_violin(position = "dodge", width = 0.8, aes(color = Outcome, fill = Outcome), alpha = 0.2) +
            geom_boxplot(position = "dodge", aes(colour = Outcome), fill = "white",  width = 0.3) +
            theme(legend.position = "none") +
            ylab("Rating")
    plot3 <- plot_grid(plot1, plot2, rel_widths = c(2, 1), labels = "AUTO")

    title <- ggdraw() + 
        draw_label(gsub("[.]", " ", var), fontface = 'bold')
    
    plot <- plot_grid(title, plot3, ncol = 1, rel_heights = c(0.1, 1))

    suppressMessages(ggsave(paste("results/dtq/", 
                    gsub("[.]", "-", var), ".png", sep = ""), plot3))
    
    plot
}
lapply(feat_lr, plot_items)

```

# Multivariate Model

## Preprocessing: Normalization + PCA
Only variables that had a pvalue < 0.05 in univariate logistic regression analysis. They were normalized and KNN imputation was used to replace missing values. 
Principal Component Analysis (PCA) was used to combine the original questionnaire items to derive features that capture the variance dataset and are uncorrelated (orthogonal).
The Principal Components that captured up to 95% of the variance of the original questionnaire data were used as predictors to estimate Outcome using Logistic Regression.

## Recursive Feature Elimination (RFE) + Multivariate Logistic Regression (LR)
Backwards Feature Selection is used to eliminate the least predictive Principal Components. 
It estimates the importance of each principal component by evaluating the difference in performance caused by the removal of individual components. 
Leave One Subject Out Cross Validation divides these datasets into train sets for modelling and test sets for estimating performance.
The AUC-ROC results achieved on the test sets by each of the models trained are considered when choosing which features to keep and remove in order to derive an optimal model.
This methodology was implemented in R using [rfe](https://search.r-project.org/CRAN/refmans/caret/html/rfe.html) (recursive feature elimination) function with ROC and [rfeControl](https://search.r-project.org/CRAN/refmans/caret/html/rfeControl.html) function with LOOCV in caret.

[Example with RF](https://towardsdatascience.com/effective-feature-selection-recursive-feature-elimination-using-r-148ff998e4f7), [ROC as metric](https://stackoverflow.com/questions/18242692/r-package-caret-rfe-function-how-to-customize-metric-to-use-auc).

```{r definition, echo = FALSE}
impute_pca <- function(df){

    print("Dimensions of original dataset:")
    print(dim(df))

    model_pca <- recipe(formula =  ~ ., data = df) %>%
                    step_normalize(all_numeric()) %>%
                    # step_impute_knn(all_numeric(), neighbors = 3) %>%
                    step_pca(all_numeric(), threshold = .95)

    model_prep <- prep(model_pca, training = df)
    return(model_prep)
}

rfe_lr <- function(df){
    #  trainctrl <- trainControl(classProbs = TRUE,
    #                        summaryFunction = twoClassSummary)
    # Setting ROC as the metric for the Logistic Regression function
    lrFuncs$summary <- twoClassSummary
    set.seed(77)    

    ctrl <- rfeControl(functions = lrFuncs,     # logistic regression
                        method = "LOOCV",       # Leave One Out Cross Validation
                        verbose = FALSE)

    # Recursive Feature Elimination with feat_lr
    model <- rfe(Outcome ~ . ,                  # predict Outcome using all other variables
                    data = df,                  # selecting the features from univariate lr
                    sizes = c(1:length(df)),
                    rfeControl = ctrl,
                    metric = "ROC") 
    
    # Printing outputs
    warnings()
    print(model)
    print(summary(model$fit))
    
    return(model)
}

dev_null_residual <- function(model){
    dev = model$null.deviance - model$deviance
    deg = model$df.null - model$df.residual
    cat('\ndeviance difference: ', dev)
    cat('\ndf difference: ', deg)
    cat('\nlevel of significance: ', pchisq(dev, deg, lower.tail = FALSE))
    cat('\nthe model is a good fit: ', pchisq(dev, deg, lower.tail = FALSE) < 0.05)
}

```

```{r, echo = FALSE}
data_prep <- impute_pca(data_lr)
data_pca  <- bake(data_prep, data_lr)
model_rfe <- rfe_lr(data_pca)
dev_null_residual(summary(model_rfe$fit))

# plot roc per rfe iteration
plot <- ggplot(data = model_rfe, metric = "ROC") + theme_bw()
suppressMessages(ggsave("results/dtq/rfe-roc.png", plot))
plot
```

The best Logistic Regression model is used to investigate the effect of the principal components on the training outcome.
The estimates were exponentiated to get the Odds Ratio (OR). The 95% Confident Interval (CI) of the ORs was also calculated. 
RFE chose 2 principal components, PC01 and PC09. 
The OR for PC01 is 2.54, this means that the probability of withdrawal increases by (2.54-1 = 1.54) 154% for every unit increase of PC1.
We are 95% confident that a unit increase in PC01 results in a 69 to 340% (1.69-1 = 0.69, 4.40-1 = 3.40) more odds of being withdrawn from training.
The OR for PC09 is 0.20, this means that the probability of withdrawal decreases by (0.20-1 = 0.80) 80% for every unit increase of PC9.
We are 95% confident that a unit increase in PC09 results in a 23 to 95% (0.04-1 = 0.96, 0.77-1 = 0.23) less odds of being withdrawn from training.

<!-- 
RFE chose 6 variables when optimizing for AUC-ROC, namely:
"Training.Trainable", "Amicability.Friendly", "Motivation.Persevering", "Training.Biddable", "Extraversion.Hyperactive", "Amicability.Easy.going".
Analysing deviance of full model compared to the null model. 
Reduction in deviance must be significant (chi-squared test) for the model to be considered a good fit. -->


## Interpretation
The best Logistic Regression model is used to investigate the effect of each variable on the training outcome.

<!-- 
Three variables showed some level of statistical significance, namely:
- "Training.Trainable" (p<0.05), for each unit increase the OR was smaller by 0.228 (less likely to be withdrawn)
- "Motivation.Persevering" (p<0.01), each unit increase the OR was smaller by 0.418 (less likely to be withdrawn)
- "Amicability.Easy.going" (p<0.1), each unit increase the OR was smaller by 0.531 (less likely to be withdrawn) -->

```{r, echo = FALSE}
results_rfe <- as.data.frame(coef(summary(model_rfe$fit)))
results_rfe$OR <- exp(results_rfe$Estimate)
results_rfe <- cbind(results_rfe, exp(confint(model_rfe$fit, level = 0.95)))
print("Results from the model created")
print(results_rfe)
```


```{r, echo = FALSE}
coef <- predictors(model_rfe)

perf <- as.data.frame(data_pca %>% select(coef,Outcome))
perf$Outcome <- relevel(perf$Outcome, "Fail")
perf$Probability <- predict(model_rfe$fit,       # model
                newdata = data_pca,        # dataframe, this will only select PC01
                type  = "response")
perf$Status <- dtq$Status
perf <- perf %>% mutate(Status = Status %>% fct_relevel(c("W", "GD", "AD")))
perf$Predicted <- ifelse(perf$Probability > 0.5, "Fail", "Success")
perf$Estimate <- log( 1 / (1/perf$Probability - 1) )

# logit function, calculate y given the probability(x)
y <- function(x){ return(log(1/( 1/x - 1))) }
th_fail = 0.65
th_success = 0.35

print("Probability predicted by the LR using PRQ data vs true Outcome")
plot <- ggplot(perf, aes(x = Estimate, y = Probability, color = Outcome )) +
    # bottom - green
    geom_rect(aes(xmin = -Inf, xmax = y(th_success), ymin = 0, ymax = th_success), 
            fill = "#99ff99", alpha = 0.2, color = "white") + 
    # top - orange
    geom_rect(aes(xmin = y(th_fail), xmax = Inf, ymin = th_fail, ymax = 1), 
            fill = "#ffff99", alpha = 0.2, color = "white") +  
    geom_point() 
plot

suppressMessages(ggsave(paste("results/dtq/logit-outcome.png"), plot))


print("Probability predicted by the LR using PRQ data vs true Status")
# plot Logit vs Status
plot <- ggplot(perf, aes(x = Estimate, y = Probability, color = Status )) +
    # bottom - green
    geom_rect(aes(xmin = -Inf, xmax = y(th_success), ymin = 0, ymax =  th_success), 
            fill = "#99ff99", alpha = 0.2, color = "white") + 
    # top - orange
    geom_rect(aes(xmin = y(th_fail), xmax = Inf, ymin =  th_fail, ymax = 1), 
            fill = "#ffff99", alpha = 0.2, color = "white") +  
    geom_point()
plot

suppressMessages(ggsave(paste("results/dtq/logit-status.png"), plot))

# calculate accuracy
sensitivity = 0.9423 # TPR -> true fail flagged as fail
specificity = 0.3889 # TNR -> true success flagged as success 
prevalence = 17/89
accuracy = (sensitivity) * (prevalence) + (specificity) * (1 - prevalence)
cat('Accuracy:', accuracy)


print("Estimate per training outcome status")
perf %>% group_by(Status) %>% dplyr::summarize(Mean = mean(Estimate, na.rm=TRUE))

cat("Predicted probability <", th_success, " -> Green flag ")
print("Dogs flagged green are likely to SUCCEED, true outcome count and status proportion:")
table(perf %>% filter(Probability < th_success) %>% pull(Outcome))
prop.table(table(perf %>% filter(Probability < th_success) %>% pull(Status)))

cat("Predicted probability >", th_fail, "-> Yellow flag")
print("Dogs flagged green are likely to FAIL, true outcome count and status proportion:")
table(perf %>% filter(Probability > th_fail) %>% pull(Outcome))
prop.table(table(perf %>% filter(Probability > th_fail) %>% pull(Status)))

```
## Feature Importance

### Explained Variance

```{r, echo = FALSE}
# tidy(recipe steps, number = select pca step, type = coef or variance
dc_pca_var <- as.data.frame(tidy(data_prep, number = 2, type = "variance"))
dc_pca_var <- dc_pca_var %>% spread(key = "terms", value = value) %>% select(-c(id))
dc_pca_var$component <- mapvalues(dc_pca_var$component, from = 1:13, 
    to = c('PC01', 'PC02', 'PC03', 'PC04', 'PC05', 'PC06', 'PC07', 'PC08', 'PC09', 'PC10', 'PC11', 'PC12', 'PC13'))
colnames(dc_pca_var) <- make.names(names(dc_pca_var)) # fixing column names with space

cat("PCA features kept in the final model after RFE: ", coef)
print("Explained Variance per Principal Component (Scree Plot)")
dc_pca_var$Included = factor(c("Yes", "No", "No", "No", "No", "No", "No", "No", "Yes", "No", "No", "No", "No"))
plot <- ggplot(dc_pca_var, aes(x = component, y = percent.variance, fill = Included)) +
    geom_bar(stat='identity', position = 'dodge', width = 0.8)+
    xlab("Principal components") +
    ylab("Explained Variance (%)") 
plot
suppressMessages(ggsave(paste("results/dtq/pc-explained-variance.png"), plot))

print("Cumulative Explained Variance per Principal Component")
plot <- ggplot(dc_pca_var, aes(x = component, y = cumulative.percent.variance, fill = Included)) +
    geom_bar(stat='identity', position = 'dodge', width = 0.8)+
    xlab("Principal Components") +
    ylab("Cumulative Variance (%)")
plot
suppressMessages(ggsave(paste("results/dtq/pc-cumulative-variance.png"), plot))

```

### Coefficient Loadings
PC01: an unit increase results in an INCREASE probability of withdrawal by 154% 
- Training-Trainable -0.3483274
- Training-Reliable -0.3388923
- Neuroticism-Fearful  0.2990981
- Amicability-Friendly -0.2980233
- Amicability-Sociable -0.2947486

PC09: an unit increase results in a DECREASED probability of withdrawal by 80%
- Training.Obedient -0.54724378 (n)
- Amicability.Easy.going  0.43676304 (n)
- Training.Trainable  0.41639880 
- Amicability.Relaxed -0.36376041  (n)
- Amicability.Sociable -0.23742114

```{r, echo = FALSE}
# model prep is the recipe steps, select pca -> number = 2, because it's the 3rd step
df_pca_coef <- as.data.frame(tidy(data_prep, number = 2, type = "coef"))
df_pca_coef$component <- mapvalues(df_pca_coef$component, 
    from = c('PC1', 'PC2', 'PC3', 'PC4', 'PC5', 'PC6', 'PC7', 'PC8', 'PC9', 'PC10', 'PC11', 'PC12', 'PC13'), 
    to = c('PC01', 'PC02', 'PC03', 'PC04', 'PC05', 'PC06', 'PC07', 'PC08', 'PC09', 'PC10', 'PC11', 'PC12', 'PC13'))
# creating another dataframe with PC1 values for the analysis
df_pca_coef <- df_pca_coef  %>% filter(grepl(paste(coef, collapse = '|'), component)) %>% select(-id)

# print the PCs values
df_pca_coef %>% 
    filter(grepl(paste(coef, collapse = '|'), component)) %>% 
    pivot_wider(names_from = component, values_from = value)

df_pca_coef %>% filter(component == 'PC01') %>% arrange(desc(abs(value))) %>% select(terms, value)
df_pca_coef %>% filter(component == 'PC09') %>% arrange(desc(abs(value))) %>% select(terms, value)

cat("Principal Component", coef," vs Original Variables")
plot <- ggplot(df_pca_coef, aes(x=terms, y=value, fill = component)) +
    geom_bar(stat="identity", position='dodge', width = 0.8) +
    xlab("MCPQ-R items - Predictor Variables") +
    ylab("Loadings") +
    theme(axis.text.x = element_text(angle = 90), legend.position = "top")
suppressMessages(ggsave(paste("results/dtq/pc-loadings.png"), plot))
plot
    
```

# Conclusion
Completing this questionnaire during training (Week 10) would allow assistance dog training organisations to understand which dog are more suitable and allow them to make informed decisions when analysing which dogs to keep for training considering the results of this objective assessment. 