---
title: "Dog Trainer Questionnaire vs Status (W, AD, GD)"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_questionnaires")
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggpubr")
library(DescTools)
library(ggplot2)
library(cowplot)
library(dplyr)
library(plyr)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(dunn.test)
```

## Introduction 
This document tests the hypothesis that there is a difference in behaviours reported in the standardised questionnaire Monash Canine Personality Questionnaire - Revised considering dog's training outcome status.

## Data Exploration
### Columns: Variables
Below are the questionnaire items(variables) that were scored by the Trainer in a scale from 1 to 6. 

```{r, echo = FALSE}
dtq = read.csv('data//2022-05-10-DT_MCPQ-R.csv', stringsAsFactors=FALSE)
vars = read.csv("data//questionnaire-variables-trainers.csv",stringsAsFactors=FALSE)

print(vars %>% filter(Type != 'info') %>% pull('Variable'))
# print the number of variables per data type
print(table(vars$Type))

# Adding a column to df vars for Test 
vars = vars %>% mutate(Test = case_when(Type == "ordinal" ~ "Kruskal"))
```


### Rows: Dogs
The number of dogs in the Dog Trainer Questionnaire dataset.
```{r, echo = FALSE}
cat('Number of dogs:', length(dtq$Code))
dtq = dtq %>% distinct(Code, .keep_all = TRUE)
cat('Number of unique dogs:', length(dtq$Code))
cat('Number of dogs per Training Outcome Status:')
table(dtq$Status)
dtq = dtq %>% filter(Status != 'in Training')
table(dtq$Status)
cat('Number of unique dogs with defined Training Outcome:', length(dtq$Code))
```

## Data Analysis
Kruskal-Wallis test was used to investigate whether there was a statistically significant difference between the behaviours reported by trainers and the training outcome status of the dogs.
The variables that showed p_value < 0.05 are below: 
```{r, echo = FALSE}
stats <- function(test, var, data, group){
  if (!is.na(test)){
    if (test == 'Kruskal'){
      #cat('Kruskal\n\n')
      r = kruskal.test(data[,var], data[,group])
      return(r$p.value)
    }
    return(NA)
  }
  return(NA)
}

vars = vars %>% mutate(status_p = round(mapply(FUN = stats, test = vars$Test, var = vars$Variable, MoreArgs = list(dtq, 'Status')),4))
vars$status_sign = ifelse(vars$status_p<0.05, 1, 0)

print(vars %>% filter(status_sign == 1) %>% select(Variable, status_p))
```

Dunn Test with the variables that showed statistical significant differences using Kruskal-Wallis.
```{r, echo = FALSE}
posthoc <- function(test, var, data, group){
  if (!is.na(test)){
    if (test == 1){
        capture.output(r <- dunn.test(data[,var], data[,group]))
        # print(var)
        # print(r$comparisons)
        # print(r$P)
        return(r$P)
      }
    return(c(NA, NA, NA))
  }
  return(c(NA, NA, NA))
}

vars = vars %>% cbind(round(matrix(mapply(FUN = posthoc, test = vars$status_sign, var = vars$Variable, MoreArgs = list(dtq, 'Status')),
                  ncol = 3, byrow = TRUE), 4))
names(vars)[6:8] <- c("AD_GD", "AD_W", "GD_W")

print(vars %>% filter(status_sign == 1) %>% select(Variable, AD_GD, AD_W, GD_W))                
```


## Training Outcome Status
Below are the variables that showed statistically significant differences in personality traits as described by their trainers considering different training outcome status ('W', 'AD', 'GD').

```{r, echo = FALSE}


vars_status =  vars %>% filter( status_sign == 1 ) %>% pull('Variable')

plotseries <- function(var){

    row = vars %>% filter(Variable == var ) %>% select("AD_GD", "AD_W", "GD_W")
    cat(var, colnames(row)[which(row < 0.05)])
    # print(row)

    plot1 <- ggplot(dtq, aes_string( x = var, fill = 'Status')) + 
            geom_bar(position = "dodge") +
            theme(legend.position = c(0.13,0.88)) +
            scale_x_discrete("Rating", limits = c(1:6), breaks = c(1:6), labels = c(1:6)) +
            expand_limits(x = c(1,6))

    plot2 <- ggplot(dtq, aes_string( x = 'Status', y = var)) + 
            geom_boxplot(position = "dodge", aes(colour = Status))  +
            theme(legend.position = "none") +
            scale_y_discrete("Rating", limits = c(1:6), breaks = c(1:6), labels = c(1:6)) +
            expand_limits(y = c(1,6)) +
            ylab("Rating")
    plot3 <- plot_grid(plot1, plot2, rel_widths = c(2, 1), labels = "AUTO")
    
    title <- ggdraw() + 
      draw_label(gsub("[.]", " - ", var), fontface = 'bold')
    
    plot <-plot_grid(title, plot3, ncol = 1, rel_heights = c(0.1, 1))
    
    suppressMessages(ggsave(paste("output/dt-analysis/3-", gsub("[.]", "-", var), ".png", sep = "") 
          , plot))
          
    plot
}


lapply(vars_status, plotseries)

```
