---
title: "Puppy Raiser Questionnaire vs Outcome (W, AD, GD)"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_questionnaires")
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggpubr")
library(DescTools)
library(ggplot2)
library(cowplot)
library(dplyr)
library(plyr)
library(tidyverse)
library(rstatix)
library(ggpubr)
```

## Introduction 
This document tests the hypothesis that there is a difference in behaviours reported in the standardised questionnaire Canine Behavioural Assessement & Research Questionnaire (C-BARQ) considering dog's training status (W, AD, GD).

## Data Exploration
### Columns: Variables
Below are the questionnaire items(variables) that were scored by the Puppy Raise in a scale from 0 = 'Never', 1 = 'Seldom', 2 = 'Sometimes', 3 = 'Usually', 4 = 'Always'. 

```{r, echo = FALSE}
prq = read.csv('data//2022-05-10-PR_C-BARQ.csv', stringsAsFactors = FALSE)
prq$Timestamp = as.Date(prq$Timestamp, format= "%Y-%m-%d")
prq$DOB = as.Date(prq$DOB, format= "%Y-%m-%d")
prq$DOA = as.Date(prq$DOA, format= "%Y-%m-%d")
prq$End.Date = as.Date(prq$End.Date, format= "%Y-%m-%d")

vars = read.csv("data//questionnaire-variables-puppyraisers.csv",stringsAsFactors=FALSE)
# print the number of variables per data type
print(table(vars$Type))

# Adding a column to df vars for Test
vars = vars %>% mutate(Test = case_when(Type == "ordinal" ~ "Kruskal"))
```


### Rows: Dogs
The number of dogs in the Puppy Raiser Questionnaire dataset. Selection criteria: dogs with a defined training outcome.
```{r, echo = FALSE}
#cat('Number of dogs:', length(prq$Name))
prq = prq %>% distinct(Code, .keep_all = TRUE)
cat('Number of unique dogs:', length(prq$Name))

cat('Training Outcome Status:')
table(prq$Status)
# filtering out dogs that don't have a defined training outcome
prq = prq %>% filter(Status != 'in Training')
cat('Number of unique dogs with defined Training Outcome:', length(prq$Name))
# filtering out dogs that don't have a defined training outcome
# prq = prq %>% filter(Status != 'CD')
```
Analysing categorical demographic data: Sex, Breed
```{r, echo = FALSE}
table(prq$Sex)
table(prq$Breed)
```
Analysing age at assessment:
```{r, echo = FALSE}
prq$Age.at.Assessment <- prq$Timestamp - prq$DOB
cat('Age at Assessment: Mean', mean(prq$Age.at.Assessment)/30.417, 'Standard Deviation', sd(prq$Age.at.Assessment)/30.417)
```
## Data 

1.	Stranger-directed aggression: Dog shows threatening or aggressive responses to strangers approaching or invading the dog’s or the owner’s personal space, territory, or home range.
2.	Owner-directed aggression: Dog shows threatening or aggressive responses to the owner or other members of the household when challenged, manhandled, stared at, stepped over, or when approached while in possession of food or objects.
3.	Dog-directed aggression: Dog shows threatening or aggressive responses when approached directly by unfamiliar dogs.
4.	Dog rivalry: Dog shows aggressive or threatening responses to other familiar dogs in the same household.
5.	Stranger-directed fear: Dog shows fearful or wary responses when approached directly by strangers.
6.	Nonsocial fear: Dog shows fearful or wary responses to sudden or loud noises (e.g. thunder), traffic, and unfamiliar objects and situations.
7.	Dog-directed fear: Dog shows fearful or wary responses when approached directly by unfamiliar dogs.
8.	Touch sensitivity: Dog shows fearful or wary responses to potentially painful or uncomfortable procedures, including bathing, grooming, nail-clipping, and veterinary examinations.  
9.	Separation-related behavior: Dog vocalizes and/or is destructive when separated from the owner, often accompanied or preceded by behavioral and autonomic signs of anxiety including restlessness, loss of appetite, trembling, and excessive salivation.
10.	Attachment and attention-seeking: Dog maintains close proximity to the owner or other members of the household, solicits affection or attention, and displays agitation when the owner gives attention to third parties.
11.	Trainability: Dog shows a willingness to attend to the owner and obey simple commands. Dog is not easily distracted, tends to be a fast learner, responds positively to correction, and will fetch or retrieve objects.
12.	Chasing: Dog chases cats, birds, and/or other small animals, given the opportunity.
13.	Excitability: Dog displays strong reaction to potentially exciting or arousing events, such as going for walks or car trips, doorbells, arrival of visitors, and the owner arriving home; has difficulty calming down after such events.
14.	Energy level: Dog is energetic, 'always on the go', and/or playful.


## Data Analysis
Kruskal-Wallis one-way analysis of variance test was used to investigate whether there was a statistically significant difference between the behaviours reported by Trainers and the Training Outcome of the dogs.
C-BARQ Correlation with Training Outcome - >Lizzy used an unequal variance t-test or Welch’s t-test was used to identify any significance
The variables that showed p_value < 0.05 are below: 
```{r, echo = FALSE}
stats <- function(test, var, data, group){
  #print(var)
  if (!is.na(test)){
    if (test == 'Kruskal'){
      #cat('Kruskal\n\n')
      r = kruskal.test(data[,var], data[,group])
      return(r$p.value)
    }
    return(NA)
  }
  return(NA)
}

vars = vars %>% mutate(status_p = mapply(FUN = stats, test = vars$Test, var = vars$Variable, MoreArgs = list(prq, 'Status')))
vars$status_sign = ifelse(vars$status_p<0.05, 1, 0)
vars %>% filter(status_sign == 1) %>% pull(Variable)
```


## Outcome
Below are the variables that showed statistically significant differences in behaviours considering different Training Outcomes (W, AD, GD).

```{r, echo = FALSE, warning = FALSE}
vars_status =  vars %>% filter( status_sign == 1 ) %>% pull('Variable')

# plotseries <- function(var){
#     ggplot(prq, aes_string(x =  'Status', fill = var)) + 
#             geom_bar(position = position_dodge2(width = 0.9, preserve = "single"))
# }
# lapply(vars_status, plotseries)



# plot_categorical <- function(var){
#     ggplot(prq, aes_string( x = var, fill = 'Status')) +
#             geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) + 
#             scale_x_discrete("Rating", limits = c(1:6), breaks = c(1:6), labels = c(1:6)) +
#             expand_limits(y = c(1,6)) +
#             ggtitle(gsub("[.]", " - ", var))
      
#   }
# lapply(vars_status[!str_detect(vars_status, "F_")][1], plot_categorical )


# plot_continuous <- function(var){
#     ggplot(prq, aes_string( x = 'Status', y = var)) + 
#     geom_boxplot(position = "dodge", aes(colour = Status)) +
#     scale_x_discrete("Rating", limits = c(1:6), breaks = c(1:6), labels = c(1:6))  +
#     expand_limits(y = c(1,6)) +
#     ggtitle(gsub("[.]", " - ", var))

#   }

# lapply(vars_status[str_detect(vars_status, "F_")], plot_continuous)


plot_categorical <- function(var){

    plot1 <- ggplot(prq, aes_string( x = var, fill = 'Status')) + 
            geom_bar(position = "dodge", width = 0.8) +
            theme(legend.position = c(0.13,0.88)) +
            scale_x_discrete("Rating", limits = c(0, 1, 2, 3, 4), breaks = c(0, 1, 2, 3, 4), labels = c('Never', 'Seldom', 'Sometimes', 'Usually', 'Always')) +
            expand_limits(x = c(0,4))

    plot2 <- ggplot(prq, aes_string( x = 'Status', y = var)) + 
            geom_boxplot(position = "dodge", aes(colour = Status))  +
            theme(legend.position = "none") +
            scale_y_discrete("Rating", limits = c(0, 1, 2, 3, 4), breaks = c(0, 1, 2, 3, 4), labels = c('Never', 'Seldom', 'Sometimes', 'Usually', 'Always')) +
            expand_limits(y = c(0,4)) +
            ylab("Rating")
    plot3 <- plot_grid(plot1, plot2, rel_widths = c(2, 1), labels = "AUTO")
    
    title <- ggdraw() + 
      draw_label(gsub("[.]", " ", var), fontface = 'bold')
    
    plot <-plot_grid(title, plot3, ncol = 1, rel_heights = c(0.1, 1))
    
    suppressMessages(ggsave(paste("output/pr-analysis/2-", sub("[.].*",  "", var), ".png", sep = ""), plot))
          
    plot
}

lapply(vars_status[!str_detect(vars_status, "F_")], plot_categorical )

plot_continuous <- function(var){
    plot <- ggplot(prq, aes_string( x = 'Status', y = var)) +
            geom_boxplot(position = "dodge2", aes(colour = Status))
            
    suppressMessages(ggsave(paste("output/pr-analysis/3-", sub("[.].*",  "", var), ".png", sep = ""), plot))
    
    plot
  }
  
lapply(vars_status[str_detect(vars_status, "F_")], plot_continuous)

```